---
title: "NMRphasing"
author: "Aixiang Jiang"
date: "June 20, 2021"
output: md_document
---


# NMRphasing
R package for NMR data phase error correction


## I. NMRphasing installation

First of all, make sure that you have R package MassSpecWavelet, which NMRphasing depends on. Example code to install MassSpecWavelet is: 

if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install("MassSpecWavelet")


Now, let's install NMRphasing from Github

```{r eval = FALSE}
devtools :: install_github(repo = "ajiangsfu/NMRphasing",force = TRUE)  ## if you do not have old versions of NMRphasing, please remove force = TRUE
```

## II. Data format

The input data format should be in one of the four formats:
1) a vector of absorption spectrum
2) a complex vector
3) a data matrix with two columns of spectrum data, the 1st column is for absorption spectrum, and 2nd column is for dispersion spectrum
4) a data frame with two columns of spectrum data, the 1st column is for absorption spectrum, and 2nd column is for dispersion spectrum

An example data can be loaded after installing NMRphasing:
```{r}
library(NMRphasing)
load(system.file("extdata", "fdat.rda", package = "NMRphasing"))
str(fdat)
```

Here, fdat$frequency_domain is the data for the following illustration, which is a complex vector. In order to make comparison, absorption part can be extracted
```{r}
Obs_absorption = Re(fdat$frequency_domain)
```



## III. Example code

There are currently seven NMR phase error correction methods implemented in this package:
"NLS", "MPC_DAOM", "MPC_EMP", "SPC_DAOM", "SPC_EMP", "SPC_AAM", and "SPC_DSM"

### 1. NLS

This is to use our new Shrinkage method to do phase error correction.
```{r}
nlsp = NMRphasing(specDatIn = fdat$frequency_domain, method = "NLS") ## NLS is the default method, method setting can be ignored

### plot to compare before and after phase error correction

nlsdat = cbind(fdat$ppm, Obs_absorption, nlsp)
nlsdat = data.frame(nlsdat)
colnames(nlsdat) = c("ppm", "Observed_Absorption", "Phased_Absoprtion")

library(ggpubr)

p1 = ggplot(nlsdat, aes(x = ppm, y = Observed_Absorption)) +
      geom_line() + theme_bw() + labs(y = "Observed Absorption") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
p2 = ggplot(nlsdat, aes(x = ppm, y = Phased_Absoprtion)) +
      geom_line() + theme_bw() + labs(y = "Phased Absoprtion") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggarrange(plotlist = list(p1,p2),labels = c("Before","After"),nrow = 2, ncol=1)

```

### 2. MPC_DAOM

This is to use our new multiple linear model approach with our new optimization function to minimize difference between absolute area and ordinary area.
```{r}
mdaomp = NMRphasing(specDatIn = fdat$frequency_domain, method = "MPC_DAOM") 
## this step might take a couple of minutes

### plot to compare before and after phase error correction

pdat = nlsdat
pdat$Phased_Absoprtion = mdaomp

p1 = ggplot(pdat, aes(x = ppm, y = Observed_Absorption)) +
      geom_line() + theme_bw() + labs(y = "Observed Absorption") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
p2 = ggplot(pdat, aes(x = ppm, y = Phased_Absoprtion)) +
      geom_line() + theme_bw() + labs(y = "Phased Absoprtion") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggarrange(plotlist = list(p1,p2),labels = c("Before","After"),nrow = 2, ncol=1)

```

### 3. MPC_EMP

This is our new multiple linear model approach based on entropy minimization with negative peak penalty
```{r}
mempp = NMRphasing(specDatIn = fdat$frequency_domain, method = "MPC_EMP") 
## this step might take a couple of minutes

### plot to compare before and after phase error correction

pdat$Phased_Absoprtion = mempp

p1 = ggplot(pdat, aes(x = ppm, y = Observed_Absorption)) +
      geom_line() + theme_bw() + labs(y = "Observed Absorption") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
p2 = ggplot(pdat, aes(x = ppm, y = Phased_Absoprtion)) +
      geom_line() + theme_bw() + labs(y = "Phased Absoprtion") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggarrange(plotlist = list(p1,p2),labels = c("Before","After"),nrow = 2, ncol=1)

```

### 4. SPC_DAOM

This phase error correction method is based on the traditional single model approach but with our new optimization function to minimize difference between absolute area and ordinal area.

```{r}
daomp = NMRphasing(specDatIn = fdat$frequency_domain, method = "SPC_DAOM") 
## this step might take a couple of minutes

### plot to compare before and after phase error correction

pdat$Phased_Absoprtion = daomp

p1 = ggplot(pdat, aes(x = ppm, y = Observed_Absorption)) +
      geom_line() + theme_bw() + labs(y = "Observed Absorption") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
p2 = ggplot(pdat, aes(x = ppm, y = Phased_Absoprtion)) +
      geom_line() + theme_bw() + labs(y = "Phased Absoprtion") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggarrange(plotlist = list(p1,p2),labels = c("Before","After"),nrow = 2, ncol=1)

```

### 5. SPC_EMP

This is an old phase error method based on the traditional single model approach targeting on entropy minimization with negative peak penalty

```{r}
empp = NMRphasing(specDatIn = fdat$frequency_domain, method = "SPC_EMP") 
## this step might take a couple of minutes

### plot to compare before and after phase error correction

pdat$Phased_Absoprtion = empp

p1 = ggplot(pdat, aes(x = ppm, y = Observed_Absorption)) +
      geom_line() + theme_bw() + labs(y = "Observed Absorption") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
p2 = ggplot(pdat, aes(x = ppm, y = Phased_Absoprtion)) +
      geom_line() + theme_bw() + labs(y = "Phased Absoprtion") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggarrange(plotlist = list(p1,p2),labels = c("Before","After"),nrow = 2, ncol=1)

```

### 6. SPC_AAM

This is an old phase error method based on the traditional single model approach with minimization on absolute area

```{r}
aamp = NMRphasing(specDatIn = fdat$frequency_domain, method = "SPC_AAM") 
## this step might take a couple of minutes

### plot to compare before and after phase error correction

pdat$Phased_Absoprtion = aamp

p1 = ggplot(pdat, aes(x = ppm, y = Observed_Absorption)) +
      geom_line() + theme_bw() + labs(y = "Observed Absorption") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
p2 = ggplot(pdat, aes(x = ppm, y = Phased_Absoprtion)) +
      geom_line() + theme_bw() + labs(y = "Phased Absoprtion") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggarrange(plotlist = list(p1,p2),labels = c("Before","After"),nrow = 2, ncol=1)

```

### 7. SPC_DSM

This is an old phase error method based on the traditional single model approach with dispersion summation minimization.

```{r}
dsmp = NMRphasing(specDatIn = fdat$frequency_domain, method = "SPC_DSM") 
## this step might take a couple of minutes

### plot to compare before and after phase error correction

pdat$Phased_Absoprtion = dsmp

p1 = ggplot(pdat, aes(x = ppm, y = Observed_Absorption)) +
      geom_line() + theme_bw() + labs(y = "Observed Absorption") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
p2 = ggplot(pdat, aes(x = ppm, y = Phased_Absoprtion)) +
      geom_line() + theme_bw() + labs(y = "Phased Absoprtion") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggarrange(plotlist = list(p1,p2),labels = c("Before","After"),nrow = 2, ncol=1)

```

## IV. NMRphasing R package general information

Version: 0.0.2

Author: Aixiang Jiang

Maintainer: Aixiang Jiang <aijiang@bccrc.ca> <aixiang.jiang@pathology.ubc.ca> 

Depends: R (>= 4.0), baseline, splines, MassSpecWavelet

Suggests: knitr

VignetteBuilder: knitr



